Software architectures for workflow integration in science gateways

### 2016-04-21 paper.tex
Added CBRAIN-FSL architecture

The architectures described hereafter also build on experience with
the SHIWA Simulation Platform~\cite{terstyanszky2014enabling}. The
SHIWA Simulation Platform is a WS-PGRADE/gUSE science gateway that was
developed in European projects
SHIWA\footnote{\url{http://www.shiwa-workflow.eu}} and
ER-flow\footnote{\url{http://www.erflow.eu}}. It focuses on workflow
interoperability for various disciplines, in particular through the
building of meta-workflows, i.e. workflows that are built out of other
workflows. It implements the so-called Coarse-Grained Interoperability
framework where workflow engines are integrated in the science gateway
as other applications and invoked through a dedicated service. The
Fine-Grained Interoperability framework was also developed in SHIWA,
where conversions between workflow languages are achieved through a
common intermediate
representation~\cite{plankensteiner-prodan-etal:2013}.

@@

The sub-tasking architecture is implemented in CBRAIN~\cite{SHER-14}
where it is used to integrate the PSOM workflow
engine~\cite{bellec2012pipeline} and the FSL toolkit. The CBRAIN-PSOM
integration~\cite{GLAT-16} is illustrated on
Figure~\ref{fig:cbrain-psom-architecture}. On this Figure, the science
gateway is represented by components \texttt{CBRAIN portal} and
\texttt{CBRAIN execution server} and the infrastructure consists of
\texttt{Storage servers} and \texttt{Computing cluster with shared
  file system}. Interaction \texttt{b} on
Figure~\ref{fig:architectures} is implemented by steps \texttt{4},
\texttt{5}, \texttt{7} and \texttt{8} (regular interactions with batch
manager). Interaction \texttt{c} is implemented by steps \texttt{3}
and \texttt{10}. Interaction \texttt{e} is implemented by step
\texttt{6}. The PSOM workflow engine adopts a pilot-job
architecture~\cite{turilli2015comprehensive} where the a master
coordinates workflow execution by submitting workers and establishing
direct communication channels with them. Note how this peculiar
execution model is totally supported by the sub-tasking architecture.

The CBRAIN-FSL integration is also remarkable as it allows to leverage
FSL pipelines that are written in a very low-level workflow language
(Linux executables and script) that submits tasks uniformly through a
specific tool called fsl\_sub. To integrate FSL in CBRAIN,
fsl\_sub was modified to implement interaction
\texttt{e}\footnote{See modified fsl\_sub script available at
  \url{https://github.com/aces/cbrain-plugins-neuro}} \todo{Add a
  diagram and maybe a code excerpt to illustrate that.}.

\begin{figure}
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-PSOM.pdf_tex}
\caption{Architecture of the CBRAIN-PSOM integration that illustrates
  the sub-tasking model (Figure extracted from~\cite{GLAT-16}). 1:
  User sends data and workflow execution request to storage server(s)
  and CBRAIN portal. 2: CBRAIN portal sends execution request to
  execution server on cluster. 3: Execution server transfers data from
  storage server(s). 4-5: Execution server starts PSOM master via task
  scheduler. 6: PSOM master submits PSOM workers to execution
  server. 7-8: Execution server starts PSOM workers through task
  scheduler. 9: PSOM master and workers execute workflow. 10.
  Execution server transfers results to storage server(s).}
\label{fig:cbrain-psom-architecture}
\end{figure}

>>>>

The architectures described hereafter also build on experience with
the SHIWA Simulation Platform~\cite{terstyanszky2014enabling}. The
SHIWA Simulation Platform is a WS-PGRADE/gUSE science gateway that was
developed in European projects
SHIWA\footnote{\url{http://www.shiwa-workflow.eu}} and
ER-flow\footnote{\url{http://www.erflow.eu}}. It focuses on workflow
interoperability for various disciplines, in particular through the
building of meta-workflows, i.e. workflows that are built out of other
workflows. It implements the so-called Coarse-Grained Interoperability
framework where workflow engines are integrated in the science gateway
as other applications and invoked through a dedicated service. The
Fine-Grained Interoperability framework was also developed in SHIWA,
where conversions between workflow languages are achieved through a
common intermediate
representation~\cite{plankensteiner-prodan-etal:2013}. \note{Tristan}{Silvia,
  do you think it is legitimate to refer to SHIWA like this in the
  paper, or should we involve Gabor in the writing?}

@@

The sub-tasking architecture is implemented in CBRAIN~\cite{SHER-14}
where it is used to integrate the FSL toolkit~\cite{Jenkinson2012782}
and the PSOM workflow engine~\cite{bellec2012pipeline}.

\begin{figure}
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-FSL.pdf_tex}
\caption{Architecture of the CBRAIN-FSL integration (sub-tasking). 1:
  User sends data and workflow execution request to storage server(s)
  and CBRAIN portal. 2: CBRAIN portal sends execution request to
  execution server on cluster. 3: Execution server transfers data from
  storage server(s). 4-5: Execution server starts FSL tool (pipeline)
  via task scheduler. 6: FSL pipeline submits sub-tasks execution
  server through fsl\_sub. 7-8: Execution server starts sub-tasks
  through task scheduler. 9.  Execution server transfers results to
  storage server(s).}
\label{fig:cbrain-fsl-architecture}
\end{figure}
The CBRAIN-FSL integration allows to leverage FSL pipelines that are
written in a very low-level workflow language (Linux executables and
script) that submits tasks uniformly through a specific tool called
fsl\_sub. It is represented in
Figure~\ref{fig:cbrain-fsl-architecture}.  On this Figure, the science
gateway is represented by components \texttt{CBRAIN portal} and
\texttt{CBRAIN execution server} and the infrastructure consists of
\texttt{Storage servers} and \texttt{Computing cluster with shared
  file system}. Interaction \texttt{b} on
Figure~\ref{fig:architectures} is implemented by steps \texttt{4},
\texttt{5}, \texttt{7} and \texttt{8} (regular interactions with batch
manager). Interaction \texttt{c} is implemented by steps \texttt{3}
and \texttt{9}. Interaction \texttt{e} is implemented by step
\texttt{6} through a modified version of the fsl\_sub
script\footnote{See modified fsl\_sub script available at
  \url{https://github.com/aces/cbrain-plugins-neuro}}.

\begin{figure}
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-PSOM.pdf_tex}
\caption{Architecture of the CBRAIN-PSOM integration (sub-tasking,
  Figure extracted from~\cite{GLAT-16}). 1: User sends data and
  workflow execution request to storage server(s) and CBRAIN
  portal. 2: CBRAIN portal sends execution request to execution server
  on cluster. 3: Execution server transfers data from storage
  server(s). 4-5: Execution server starts PSOM master via task
  scheduler. 6: PSOM master submits PSOM workers to execution
  server. 7-8: Execution server starts PSOM workers through task
  scheduler. 9: PSOM master and workers execute workflow. 10.
  Execution server transfers results to storage server(s).}
\label{fig:cbrain-psom-architecture}
\end{figure}
The CBRAIN-PSOM integration~\cite{GLAT-16} is illustrated on
Figure~\ref{fig:cbrain-psom-architecture}. The PSOM workflow engine
adopts a pilot-job architecture~\cite{turilli2015comprehensive} where
the a master coordinates workflow execution by submitting workers and
establishing direct communication channels with them. Note how this
peculiar execution model is totally supported by the sub-tasking
architecture.
